---
title: "Advanced Multi-Provider Features"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Advanced Multi-Provider Features}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Advanced Multi-Provider Features

This guide covers advanced features of the multi-provider mapdeck system, including provider switching, coordinate transformations, performance optimization, and integration patterns.

## Provider Switching

### Dynamic Provider Switching

Switch between providers while preserving your visualization:

```{r}
library(mapdeck)

# Create initial map with layers
original_map <- mapdeck(
  provider = "mapbox",
  style = "streets",
  location = c(-74.006, 40.7128),
  zoom = 12
) %>%
  add_scatterplot(
    data = capitals,
    lon = "lon",
    lat = "lat",
    radius = 50000,
    fill_colour = "country",
    tooltip = c("capital_city", "country")
  ) %>%
  add_polygon(
    data = melbourne,
    fill_colour = "SA2_NAME",
    fill_opacity = 0.5,
    layer_id = "boundaries"
  )

# Switch to Leaflet while preserving all layers
leaflet_map <- update_provider(original_map, "leaflet")

# Switch to OpenLayers with custom configuration
openlayers_map <- update_provider(
  original_map,
  "openlayers",
  config = list(
    projection = "EPSG:3857",
    source_type = "OSM"
  )
)
```

### Compatibility Checking

Check compatibility before switching providers:

```{r}
# Check compatibility between providers
compatibility <- check_provider_compatibility("mapbox", "leaflet")

print("Compatibility Results:")
print(paste("Compatible:", compatibility$compatible))
print(paste("Reason:", compatibility$reason))

if (length(compatibility$issues) > 0) {
  print("Issues:")
  for (issue in compatibility$issues) {
    print(paste("  -", issue))
  }
}

if (length(compatibility$degraded_features) > 0) {
  print("Features that will be degraded:")
  for (feature in compatibility$degraded_features) {
    print(paste("  -", feature))
  }
}

# Check coordinate transformation requirements
print(paste("Coordinate transformation required:", 
           compatibility$coordinate_transformation_required))
```

### Graceful Degradation

Handle provider limitations automatically:

```{r}
# Create 3D visualization with Mapbox
mapbox_3d <- mapdeck(
  provider = "mapbox",
  style = "dark",
  location = c(-122.4194, 37.7749),
  zoom = 15,
  pitch = 60,
  bearing = 45
) %>%
  add_column(
    data = capitals,
    lon = "lon",
    lat = "lat",
    elevation = "population",
    fill_colour = "country",
    radius = 10000
  )

# Switch to Leaflet - 3D features are gracefully degraded
leaflet_2d <- update_provider(
  mapbox_3d,
  "leaflet",
  preserve_layers = TRUE,
  graceful_degradation = TRUE
)

# The column layer becomes a scaled scatterplot layer
# Pitch and bearing are set to 0 (with warnings)
```

## Coordinate System Transformations

### Automatic Coordinate Handling

The system automatically handles coordinate transformations:

```{r}
# WGS84 data (standard GPS coordinates)
wgs84_data <- data.frame(
  location = c("Beijing", "Shanghai", "Guangzhou"),
  lon = c(116.3974, 121.4737, 113.2644),
  lat = c(39.9093, 31.2304, 23.1291),
  value = c(100, 150, 120)
)

# Mapbox uses WGS84 (no transformation needed)
mapbox_map <- mapdeck(provider = "mapbox") %>%
  add_scatterplot(data = wgs84_data, lon = "lon", lat = "lat")

# Gaode uses GCJ02 (automatic transformation)
gaode_map <- mapdeck(provider = "gaode") %>%
  add_scatterplot(data = wgs84_data, lon = "lon", lat = "lat")
# Coordinates automatically transformed from WGS84 to GCJ02

# Baidu uses BD09 (automatic transformation)
baidu_map <- mapdeck(provider = "baidu") %>%
  add_scatterplot(data = wgs84_data, lon = "lon", lat = "lat")
# Coordinates automatically transformed from WGS84 to BD09
```

### Manual Coordinate Transformations

For advanced use cases, perform manual transformations:

```{r}
# Manual coordinate transformation
wgs84_coords <- data.frame(
  lon = c(116.3974, 121.4737),
  lat = c(39.9093, 31.2304)
)

# Transform to different coordinate systems
gcj02_coords <- transform_coordinates(wgs84_coords, "WGS84", "GCJ02")
bd09_coords <- transform_coordinates(wgs84_coords, "WGS84", "BD09")

print("Original WGS84:")
print(wgs84_coords)
print("Transformed to GCJ02 (Gaode):")
print(gcj02_coords)
print("Transformed to BD09 (Baidu):")
print(bd09_coords)

# Detect coordinate system automatically
detected_crs <- detect_coordinate_system(wgs84_coords)
print(paste("Detected coordinate system:", detected_crs))

# Transform data coordinates for specific providers
transformed_data <- transform_data_coordinates(
  wgs84_data, 
  from_crs = "WGS84", 
  to_crs = "GCJ02"
)
```

### Coordinate Accuracy Validation

Validate transformation accuracy:

```{r}
# Test coordinate transformation accuracy
test_accuracy <- function(original_coords, target_crs) {
  # Transform coordinates
  transformed <- transform_coordinates(original_coords, "WGS84", target_crs)
  
  # Transform back to WGS84
  back_transformed <- transform_coordinates(transformed, target_crs, "WGS84")
  
  # Calculate accuracy (in meters)
  lon_diff <- original_coords$lon - back_transformed$lon
  lat_diff <- original_coords$lat - back_transformed$lat
  
  # Convert to meters (approximate)
  accuracy_meters <- sqrt(lon_diff^2 + lat_diff^2) * 111000
  
  return(list(
    original = original_coords,
    transformed = transformed,
    back_transformed = back_transformed,
    accuracy_meters = accuracy_meters
  ))
}

# Test accuracy for different coordinate systems
beijing_coords <- data.frame(lon = 116.3974, lat = 39.9093)

gcj02_accuracy <- test_accuracy(beijing_coords, "GCJ02")
bd09_accuracy <- test_accuracy(beijing_coords, "BD09")

print(paste("GCJ02 transformation accuracy:", 
           round(gcj02_accuracy$accuracy_meters, 2), "meters"))
print(paste("BD09 transformation accuracy:", 
           round(bd09_accuracy$accuracy_meters, 2), "meters"))
```

## Advanced Styling and Theming

### Cross-Provider Style Management

Manage styles consistently across providers:

```{r}
# Get available styles for each provider
mapbox_styles <- get_available_styles("mapbox")
leaflet_styles <- get_available_styles("leaflet")
gaode_styles <- get_available_styles("gaode")

print("Mapbox styles:")
print(mapbox_styles)
print("Leaflet styles:")
print(leaflet_styles)
print("Gaode styles:")
print(gaode_styles)

# Create style mapping function
create_styled_map <- function(provider, theme = "dark") {
  
  # Map generic themes to provider-specific styles
  style_mapping <- list(
    mapbox = list(
      light = "streets",
      dark = "dark",
      satellite = "satellite"
    ),
    leaflet = list(
      light = "light",  # CartoDB Positron
      dark = "dark",    # CartoDB DarkMatter
      satellite = "satellite"  # Esri WorldImagery
    ),
    gaode = list(
      light = "normal",
      dark = "dark",
      satellite = "satellite"
    )
  )
  
  style <- style_mapping[[provider]][[theme]]
  
  mapdeck(
    provider = provider,
    style = style,
    location = c(0, 20),
    zoom = 2
  )
}

# Create consistent dark theme across providers
dark_mapbox <- create_styled_map("mapbox", "dark")
dark_leaflet <- create_styled_map("leaflet", "dark")
dark_gaode <- create_styled_map("gaode", "dark")
```

### Custom Style Validation

Validate styles before applying:

```{r}
# Validate style compatibility
validate_style <- function(style, provider) {
  validation <- validate_map_style(style, provider, detailed = TRUE)
  
  cat("Style Validation Results:\n")
  cat("Style:", style, "\n")
  cat("Provider:", provider, "\n")
  cat("Valid:", validation$valid, "\n")
  
  if (!validation$valid) {
    cat("Issues:\n")
    for (issue in validation$issues) {
      cat("  -", issue, "\n")
    }
    
    if (!is.null(validation$suggested_alternatives)) {
      cat("Suggested alternatives:\n")
      for (alt in validation$suggested_alternatives) {
        cat("  -", alt, "\n")
      }
    }
  }
  
  return(validation)
}

# Test style validation
validate_style("streets", "mapbox")
validate_style("watercolor", "leaflet")
validate_style("invalid_style", "mapbox")
```

### Theme Management

Apply consistent themes across providers:

```{r}
# Create theme manager
theme_manager <- list(
  light = list(
    background = "#FFFFFF",
    text_color = "#2C3E50",
    accent_color = "#3498DB",
    mapbox_style = "streets",
    leaflet_style = "light",
    gaode_style = "normal"
  ),
  dark = list(
    background = "#2C3E50",
    text_color = "#FFFFFF",
    accent_color = "#E74C3C",
    mapbox_style = "dark",
    leaflet_style = "dark",
    gaode_style = "dark"
  ),
  satellite = list(
    background = "#000000",
    text_color = "#FFFFFF",
    accent_color = "#F39C12",
    mapbox_style = "satellite",
    leaflet_style = "satellite",
    gaode_style = "satellite"
  )
)

# Apply theme to map
apply_theme <- function(provider, theme_name) {
  theme <- theme_manager[[theme_name]]
  style_key <- paste0(provider, "_style")
  
  map <- mapdeck(
    provider = provider,
    style = theme[[style_key]],
    location = c(0, 20),
    zoom = 2
  )
  
  # Apply theme-specific styling to layers
  map %>%
    add_scatterplot(
      data = capitals,
      lon = "lon",
      lat = "lat",
      radius = 100000,
      fill_colour = theme$accent_color,
      stroke_colour = theme$text_color
    )
}

# Create themed maps
light_mapbox <- apply_theme("mapbox", "light")
dark_leaflet <- apply_theme("leaflet", "dark")
satellite_gaode <- apply_theme("gaode", "satellite")
```

## Performance Optimization

### Provider Performance Comparison

Compare performance across providers:

```{r}
# Performance testing function
benchmark_providers <- function(data, layer_type = "scatterplot") {
  
  providers <- list_available_providers()
  results <- list()
  
  for (provider in providers) {
    cat("Benchmarking", provider, "...\n")
    
    # Skip providers that require authentication if not configured
    if (provider %in% c("mapbox", "gaode", "baidu")) {
      token <- get_access_token(provider)
      if (is.null(token) || token == "") {
        cat("  Skipped: No authentication token\n")
        next
      }
    }
    
    # Measure rendering time
    start_time <- Sys.time()
    
    tryCatch({
      map <- mapdeck(
        provider = provider,
        location = c(0, 0),
        zoom = 2
      )
      
      if (layer_type == "scatterplot") {
        map <- map %>%
          add_scatterplot(
            data = data,
            lon = "lon",
            lat = "lat",
            radius = 50000,
            fill_colour = "country"
          )
      } else if (layer_type == "screengrid") {
        map <- map %>%
          add_screengrid(
            data = data,
            lon = "lon",
            lat = "lat",
            cell_size = 50
          )
      }
      
      end_time <- Sys.time()
      results[[provider]] <- list(
        time_seconds = as.numeric(end_time - start_time),
        success = TRUE,
        error = NULL
      )
      
      cat("  Time:", round(results[[provider]]$time_seconds, 3), "seconds\n")
      
    }, error = function(e) {
      results[[provider]] <- list(
        time_seconds = NA,
        success = FALSE,
        error = e$message
      )
      cat("  Error:", e$message, "\n")
    })
  }
  
  return(results)
}

# Run performance benchmark
if (exists("capitals")) {
  performance_results <- benchmark_providers(capitals, "scatterplot")
  
  # Analyze results
  successful_results <- performance_results[
    sapply(performance_results, function(x) x$success)
  ]
  
  if (length(successful_results) > 0) {
    times <- sapply(successful_results, function(x) x$time_seconds)
    fastest_provider <- names(times)[which.min(times)]
    
    cat("\nPerformance Summary:\n")
    cat("Fastest provider:", fastest_provider, 
        "(" , round(min(times), 3), "seconds)\n")
    
    for (provider in names(times)) {
      cat(provider, ":", round(times[provider], 3), "seconds\n")
    }
  }
}
```

### Memory Usage Optimization

Monitor and optimize memory usage:

```{r}
# Memory usage monitoring
monitor_memory_usage <- function(provider, data_size = 1000) {
  
  # Generate test data
  test_data <- data.frame(
    lon = runif(data_size, -180, 180),
    lat = runif(data_size, -85, 85),
    value = rnorm(data_size, 100, 20)
  )
  
  # Monitor memory before
  gc()  # Garbage collection
  mem_before <- memory.size()
  
  # Create map with data
  map <- mapdeck(provider = provider, location = c(0, 0), zoom = 2) %>%
    add_scatterplot(
      data = test_data,
      lon = "lon",
      lat = "lat",
      radius = 10000,
      fill_colour = "value"
    )
  
  # Monitor memory after
  mem_after <- memory.size()
  memory_used <- mem_after - mem_before
  
  cat("Memory usage for", provider, "with", data_size, "points:\n")
  cat("  Memory used:", round(memory_used, 2), "MB\n")
  cat("  Memory per point:", round(memory_used / data_size * 1000, 2), "KB\n")
  
  return(list(
    provider = provider,
    data_size = data_size,
    memory_mb = memory_used,
    memory_per_point_kb = memory_used / data_size * 1000
  ))
}

# Test memory usage across providers
memory_results <- list()
data_sizes <- c(1000, 5000, 10000)

for (size in data_sizes) {
  for (provider in c("mapbox", "leaflet", "openlayers")) {
    if (provider %in% list_available_providers()) {
      result <- monitor_memory_usage(provider, size)
      memory_results[[paste(provider, size, sep = "_")]] <- result
    }
  }
}
```

### Large Dataset Handling

Optimize for large datasets:

```{r}
# Large dataset optimization strategies
optimize_large_dataset <- function(data, provider, optimization = "auto") {
  
  data_size <- nrow(data)
  cat("Optimizing", data_size, "points for", provider, "\n")
  
  # Choose optimization strategy based on data size and provider
  if (optimization == "auto") {
    if (data_size < 1000) {
      optimization <- "none"
    } else if (data_size < 10000) {
      optimization <- "sampling"
    } else {
      optimization <- "aggregation"
    }
  }
  
  cat("Using optimization strategy:", optimization, "\n")
  
  map <- mapdeck(provider = provider, location = c(0, 0), zoom = 2)
  
  if (optimization == "none") {
    # No optimization - render all points
    map <- map %>%
      add_scatterplot(
        data = data,
        lon = "lon",
        lat = "lat",
        radius = 5000,
        fill_colour = "value"
      )
      
  } else if (optimization == "sampling") {
    # Sample data to reduce point count
    sample_size <- min(5000, nrow(data))
    sampled_data <- data[sample(nrow(data), sample_size), ]
    
    map <- map %>%
      add_scatterplot(
        data = sampled_data,
        lon = "lon",
        lat = "lat",
        radius = 8000,
        fill_colour = "value"
      )
      
  } else if (optimization == "aggregation") {
    # Use aggregation layers for large datasets
    map <- map %>%
      add_screengrid(
        data = data,
        lon = "lon",
        lat = "lat",
        weight = "value",
        cell_size = 50,
        opacity = 0.8
      )
  }
  
  return(map)
}

# Test with large dataset
if (FALSE) {  # Example with large dataset
  large_data <- data.frame(
    lon = runif(50000, -180, 180),
    lat = runif(50000, -85, 85),
    value = rnorm(50000, 100, 20)
  )
  
  # Test different optimization strategies
  no_opt_map <- optimize_large_dataset(large_data, "leaflet", "none")
  sampled_map <- optimize_large_dataset(large_data, "leaflet", "sampling")
  aggregated_map <- optimize_large_dataset(large_data, "leaflet", "aggregation")
}
```

## Integration Patterns

### Shiny Integration

Advanced Shiny integration with provider switching:

```{r}
if (FALSE) {  # Shiny application example
  library(shiny)
  library(shinydashboard)
  
  ui <- dashboardPage(
    dashboardHeader(title = "Multi-Provider Map Dashboard"),
    
    dashboardSidebar(
      sidebarMenu(
        menuItem("Map View", tabName = "map", icon = icon("map")),
        menuItem("Settings", tabName = "settings", icon = icon("cog"))
      )
    ),
    
    dashboardBody(
      tabItems(
        tabItem(tabName = "map",
          fluidRow(
            box(
              title = "Map Controls", status = "primary", solidHeader = TRUE,
              width = 3,
              selectInput("provider", "Map Provider:",
                         choices = list_available_providers(),
                         selected = "mapbox"),
              selectInput("style", "Map Style:",
                         choices = NULL),
              sliderInput("zoom", "Zoom Level:", 
                         min = 1, max = 18, value = 10),
              checkboxInput("enable_3d", "Enable 3D (if supported)", FALSE),
              actionButton("switch_provider", "Switch Provider", 
                          class = "btn-warning"),
              hr(),
              h4("Provider Info"),
              verbatimTextOutput("provider_capabilities")
            ),
            box(
              title = "Map", status = "primary", solidHeader = TRUE,
              width = 9,
              mapdeckOutput("map", height = "600px")
            )
          )
        ),
        
        tabItem(tabName = "settings",
          fluidRow(
            box(
              title = "Authentication", status = "primary", solidHeader = TRUE,
              width = 6,
              textInput("mapbox_token", "Mapbox Token:", 
                       placeholder = "pk.your_token_here"),
              textInput("gaode_token", "Gaode API Key:", 
                       placeholder = "your_gaode_key_here"),
              textInput("baidu_token", "Baidu API Key:", 
                       placeholder = "your_baidu_key_here"),
              actionButton("save_tokens", "Save Tokens", class = "btn-success")
            ),
            box(
              title = "Performance", status = "primary", solidHeader = TRUE,
              width = 6,
              numericInput("max_points", "Max Points to Display:", 
                          value = 5000, min = 100, max = 50000),
              selectInput("optimization", "Optimization Strategy:",
                         choices = c("Auto" = "auto", 
                                   "None" = "none",
                                   "Sampling" = "sampling", 
                                   "Aggregation" = "aggregation")),
              checkboxInput("enable_gpu", "Enable GPU Acceleration", TRUE)
            )
          )
        )
      )
    )
  )
  
  server <- function(input, output, session) {
    
    # Reactive values
    values <- reactiveValues(
      current_map = NULL,
      provider_switched = FALSE
    )
    
    # Update style choices based on provider
    observe({
      req(input$provider)
      styles <- get_available_styles(input$provider)
      updateSelectInput(session, "style", choices = styles, selected = styles[1])
    })
    
    # Save authentication tokens
    observeEvent(input$save_tokens, {
      if (input$mapbox_token != "") {
        set_token(input$mapbox_token, provider = "mapbox")
      }
      if (input$gaode_token != "") {
        set_token(input$gaode_token, provider = "gaode")
      }
      if (input$baidu_token != "") {
        set_token(input$baidu_token, provider = "baidu")
      }
      
      showNotification("Tokens saved successfully!", type = "success")
    })
    
    # Create reactive map
    map_reactive <- reactive({
      
      # Get provider capabilities
      capabilities <- get_provider_capabilities(input$provider)
      
      # Adjust 3D settings based on provider support
      pitch <- if (input$enable_3d && capabilities$supports_3d) 45 else 0
      bearing <- if (input$enable_3d && capabilities$supports_3d) 30 else 0
      
      map <- mapdeck(
        provider = input$provider,
        style = input$style,
        zoom = input$zoom,
        pitch = pitch,
        bearing = bearing,
        location = c(0, 20)
      ) %>%
        add_scatterplot(
          data = capitals,
          lon = "lon",
          lat = "lat",
          radius = 100000,
          fill_colour = "country",
          tooltip = c("capital_city", "country")
        )
      
      values$current_map <- map
      return(map)
    })
    
    # Render map
    output$map <- renderMapdeck({
      map_reactive()
    })
    
    # Provider switching
    observeEvent(input$switch_provider, {
      req(values$current_map)
      
      # Check compatibility
      current_provider <- extract_provider_info(values$current_map)$provider
      compatibility <- check_provider_compatibility(current_provider, input$provider)
      
      if (compatibility$compatible) {
        # Switch provider
        new_map <- update_provider(
          values$current_map,
          input$provider,
          preserve_layers = TRUE,
          transform_coordinates = TRUE
        )
        
        values$current_map <- new_map
        values$provider_switched <- TRUE
        
        output$map <- renderMapdeck({
          new_map
        })
        
        showNotification(
          paste("Switched to", input$provider, "successfully!"),
          type = "success"
        )
      } else {
        showNotification(
          paste("Cannot switch to", input$provider, ":", compatibility$reason),
          type = "error"
        )
      }
    })
    
    # Display provider capabilities
    output$provider_capabilities <- renderText({
      req(input$provider)
      capabilities <- get_provider_capabilities(input$provider)
      
      paste(
        "Provider:", input$provider,
        "\nSupports 3D:", capabilities$supports_3d,
        "\nCoordinate System:", capabilities$coordinate_system,
        "\nAuthentication Required:", capabilities$requires_auth,
        "\nMax Zoom Level:", capabilities$max_zoom,
        "\nSupported Layers:", paste(capabilities$supported_layers, collapse = ", ")
      )
    })
  }
  
  shinyApp(ui = ui, server = server)
}
```

### API Integration

Integrate with external APIs and data sources:

```{r}
# Example: Weather data integration
integrate_weather_data <- function(provider, api_key) {
  
  # Mock weather API call (replace with actual API)
  get_weather_data <- function() {
    data.frame(
      city = c("New York", "London", "Tokyo", "Sydney"),
      lon = c(-74.006, -0.1276, 139.6917, 151.2093),
      lat = c(40.7128, 51.5074, 35.6895, -33.8688),
      temperature = c(22, 18, 28, 25),
      humidity = c(65, 80, 70, 55),
      weather = c("Sunny", "Cloudy", "Rainy", "Clear")
    )
  }
  
  weather_data <- get_weather_data()
  
  # Create weather visualization
  weather_map <- mapdeck(
    provider = provider,
    location = c(0, 30),
    zoom = 2
  ) %>%
    add_scatterplot(
      data = weather_data,
      lon = "lon",
      lat = "lat",
      radius = "temperature",
      fill_colour = "weather",
      tooltip = c("city", "temperature", "humidity", "weather"),
      radius_min_pixels = 10,
      radius_max_pixels = 50
    ) %>%
    add_text(
      data = weather_data,
      lon = "lon",
      lat = "lat",
      text = "temperature",
      size = 12,
      colour = "#FFFFFF"
    )
  
  return(weather_map)
}

# Use with different providers
if (FALSE) {
  weather_mapbox <- integrate_weather_data("mapbox", "your_weather_api_key")
  weather_leaflet <- integrate_weather_data("leaflet", "your_weather_api_key")
}
```

This comprehensive guide covers the advanced features available in the multi-provider mapdeck system, enabling you to build sophisticated mapping applications with provider flexibility, coordinate system handling, and performance optimization.