---
title: "Migration Guide: Single to Multi-Provider Usage"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Migration Guide: Single to Multi-Provider Usage}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Migration Guide: Single to Multi-Provider Usage

This guide helps you migrate from the single-provider mapdeck to the new multi-provider version while maintaining backward compatibility and exploring new features.

## Overview

The multi-provider version of mapdeck maintains **100% backward compatibility** with existing code while adding support for multiple mapping providers. Your existing code will continue to work without any modifications.

## Backward Compatibility

### Existing Code Continues to Work

All your existing mapdeck code will work exactly as before:

```{r}
# This code works identically in both versions
library(mapdeck)

# Set token (still works for Mapbox)
set_token("pk.your_mapbox_token")

# Create map (defaults to Mapbox)
map <- mapdeck(
  location = c(-74.006, 40.7128),
  zoom = 12,
  style = "streets"
)

# Add layers (no changes)
map %>%
  add_scatterplot(
    data = capitals,
    lon = "lon",
    lat = "lat",
    radius = 50000,
    fill_colour = "country"
  )
```

### What Hasn't Changed

- All existing function signatures remain the same
- Default behavior is identical (Mapbox provider)
- All layer functions work exactly as before
- Token management for Mapbox is unchanged
- Style names and options are preserved
- Shiny integration remains the same

## New Multi-Provider Features

### 1. Provider Selection

The main new feature is the ability to specify different providers:

```{r}
# Old way (still works)
map_old <- mapdeck(location = c(0, 0), zoom = 2)

# New way with explicit provider
map_new <- mapdeck(
  provider = "mapbox",  # Explicit provider selection
  location = c(0, 0), 
  zoom = 2
)

# Alternative providers
leaflet_map <- mapdeck(provider = "leaflet", location = c(0, 0), zoom = 2)
gaode_map <- mapdeck(provider = "gaode", location = c(116.4, 39.9), zoom = 10)
```

### 2. Enhanced Token Management

Token management now supports multiple providers:

```{r}
# Old way (still works for Mapbox)
set_token("pk.your_mapbox_token")

# New way with provider specification
set_token("pk.your_mapbox_token", provider = "mapbox")
set_token("your_gaode_api_key", provider = "gaode")
set_token("your_baidu_api_key", provider = "baidu")

# View all tokens
mapdeck_tokens()
```

### 3. Provider-Specific Styling

Styles now adapt to the selected provider:

```{r}
# Mapbox styles
mapbox_map <- mapdeck(provider = "mapbox", style = "streets")

# Leaflet tile providers
leaflet_map <- mapdeck(provider = "leaflet", style = "streets")  # OSM
leaflet_dark <- mapdeck(provider = "leaflet", style = "dark")    # CartoDB Dark

# Chinese provider styles
gaode_map <- mapdeck(provider = "gaode", style = "satellite")
```

## Step-by-Step Migration

### Step 1: Update Your Package

```{r}
# Install the multi-provider version
# (Your existing code will still work)
library(mapdeck)
```

### Step 2: Test Existing Code

Run your existing code to ensure everything works:

```{r}
# Your existing code should work without changes
existing_map <- mapdeck(
  token = "pk.your_token",  # Still works
  style = "streets",
  location = c(-74.006, 40.7128),
  zoom = 12
) %>%
  add_scatterplot(
    data = your_data,
    lon = "longitude",
    lat = "latitude",
    radius = 1000,
    fill_colour = "category"
  )
```

### Step 3: Gradually Adopt New Features

Start using new features incrementally:

```{r}
# Add explicit provider specification
updated_map <- mapdeck(
  provider = "mapbox",      # NEW: Explicit provider
  token = "pk.your_token",  # Same as before
  style = "streets",
  location = c(-74.006, 40.7128),
  zoom = 12
) %>%
  add_scatterplot(
    data = your_data,
    lon = "longitude",
    lat = "latitude",
    radius = 1000,
    fill_colour = "category"
  )
```

### Step 4: Explore Alternative Providers

Try different providers for comparison:

```{r}
# Create the same visualization with different providers
create_comparison_maps <- function(data) {
  
  base_config <- list(
    location = c(-74.006, 40.7128),
    zoom = 12
  )
  
  # Mapbox version
  mapbox_map <- mapdeck(provider = "mapbox", !!!base_config) %>%
    add_scatterplot(data = data, lon = "lon", lat = "lat", 
                   radius = 1000, fill_colour = "category")
  
  # Leaflet version
  leaflet_map <- mapdeck(provider = "leaflet", !!!base_config) %>%
    add_scatterplot(data = data, lon = "lon", lat = "lat", 
                   radius = 1000, fill_colour = "category")
  
  return(list(mapbox = mapbox_map, leaflet = leaflet_map))
}
```

## Common Migration Scenarios

### Scenario 1: Basic Map Creation

**Before:**
```{r}
map <- mapdeck(
  token = "pk.your_token",
  location = c(-74.006, 40.7128),
  zoom = 12
)
```

**After (with new features):**
```{r}
# Set token globally (recommended)
set_token("pk.your_token", provider = "mapbox")

# Create map with explicit provider
map <- mapdeck(
  provider = "mapbox",
  location = c(-74.006, 40.7128),
  zoom = 12
)

# Or try alternative provider
leaflet_map <- mapdeck(
  provider = "leaflet",
  location = c(-74.006, 40.7128),
  zoom = 12
)
```

### Scenario 2: Multiple Maps in Application

**Before:**
```{r}
# Multiple maps with same provider
map1 <- mapdeck(token = "pk.token", style = "streets")
map2 <- mapdeck(token = "pk.token", style = "satellite")
map3 <- mapdeck(token = "pk.token", style = "dark")
```

**After:**
```{r}
# Set token once
set_token("pk.your_token", provider = "mapbox")

# Mix providers for different use cases
overview_map <- mapdeck(provider = "leaflet", style = "streets")    # Free overview
detail_map <- mapdeck(provider = "mapbox", style = "satellite")     # Detailed satellite
analysis_map <- mapdeck(provider = "mapbox", style = "dark")        # Dark theme for analysis
```

### Scenario 3: Shiny Applications

**Before:**
```{r}
# Shiny server function
server <- function(input, output) {
  output$map <- renderMapdeck({
    mapdeck(
      token = "pk.your_token",
      style = input$style,
      location = c(input$lon, input$lat),
      zoom = input$zoom
    )
  })
}
```

**After:**
```{r}
# Enhanced Shiny server with provider selection
server <- function(input, output) {
  output$map <- renderMapdeck({
    mapdeck(
      provider = input$provider,  # NEW: User can select provider
      style = input$style,
      location = c(input$lon, input$lat),
      zoom = input$zoom
    )
  })
}

# Enhanced UI with provider selection
ui <- fluidPage(
  selectInput("provider", "Map Provider:",
             choices = list_available_providers(),
             selected = "mapbox"),
  selectInput("style", "Map Style:",
             choices = NULL),  # Populated based on provider
  mapdeckOutput("map")
)
```

### Scenario 4: Data Analysis Workflows

**Before:**
```{r}
analyze_spatial_data <- function(data) {
  map <- mapdeck(token = "pk.token") %>%
    add_scatterplot(data = data, lon = "lon", lat = "lat")
  
  return(map)
}
```

**After:**
```{r}
analyze_spatial_data <- function(data, provider = "mapbox") {
  # Provider-agnostic analysis
  map <- mapdeck(provider = provider) %>%
    add_scatterplot(data = data, lon = "lon", lat = "lat")
  
  return(map)
}

# Use different providers for different purposes
exploratory_map <- analyze_spatial_data(data, "leaflet")    # Free exploration
publication_map <- analyze_spatial_data(data, "mapbox")     # High-quality output
```

## Advanced Migration Features

### Provider Switching

Switch between providers while preserving your visualization:

```{r}
# Create map with one provider
original_map <- mapdeck(provider = "mapbox") %>%
  add_scatterplot(data = capitals, lon = "lon", lat = "lat", 
                 radius = 50000, fill_colour = "country")

# Switch to different provider (preserves layers)
switched_map <- update_provider(original_map, "leaflet")
```

### Coordinate System Handling

The system automatically handles coordinate transformations:

```{r}
# Your WGS84 data works with all providers
wgs84_data <- data.frame(
  lon = c(116.3974, 121.4737),  # Beijing, Shanghai
  lat = c(39.9093, 31.2304)
)

# Automatically transforms to appropriate coordinate system
mapbox_map <- mapdeck(provider = "mapbox") %>%
  add_scatterplot(data = wgs84_data, lon = "lon", lat = "lat")

gaode_map <- mapdeck(provider = "gaode") %>%
  add_scatterplot(data = wgs84_data, lon = "lon", lat = "lat")  # Auto-transforms to GCJ02
```

### Style Compatibility

Styles are automatically mapped between providers:

```{r}
# Generic style names work across providers
create_styled_map <- function(provider) {
  mapdeck(provider = provider, style = "dark") %>%
    add_scatterplot(data = capitals, lon = "lon", lat = "lat")
}

mapbox_dark <- create_styled_map("mapbox")    # Uses Mapbox dark style
leaflet_dark <- create_styled_map("leaflet")  # Uses CartoDB DarkMatter
```

## Best Practices for Migration

### 1. Gradual Adoption

- Start by adding explicit `provider = "mapbox"` to existing code
- Test one provider at a time
- Gradually explore alternative providers

### 2. Token Management

```{r}
# Recommended approach
set_token("pk.mapbox_token", provider = "mapbox")
set_token("gaode_api_key", provider = "gaode")
set_token("baidu_api_key", provider = "baidu")

# Check all tokens are set
mapdeck_tokens()
```

### 3. Provider Selection Strategy

```{r}
# Choose providers based on use case
choose_provider <- function(use_case) {
  switch(use_case,
    "exploration" = "leaflet",      # Free, good for exploration
    "publication" = "mapbox",       # High quality, 3D support
    "china" = "gaode",             # Optimized for China
    "analysis" = "mapbox"          # Advanced features
  )
}

provider <- choose_provider("exploration")
map <- mapdeck(provider = provider)
```

### 4. Error Handling

```{r}
# Robust provider selection with fallbacks
create_map_safely <- function(preferred_provider = "mapbox") {
  tryCatch({
    mapdeck(provider = preferred_provider)
  }, error = function(e) {
    warning(paste("Failed to create", preferred_provider, "map:", e$message))
    # Fallback to Leaflet (no authentication required)
    mapdeck(provider = "leaflet")
  })
}
```

## Testing Your Migration

### Compatibility Check

```{r}
# Test that your existing code still works
test_backward_compatibility <- function() {
  
  # Test 1: Basic map creation
  map1 <- mapdeck(location = c(0, 0), zoom = 2)
  stopifnot(inherits(map1, "mapdeck"))
  
  # Test 2: Token setting
  set_token("test_token")
  token <- get_access_token()
  stopifnot(!is.null(token))
  
  # Test 3: Layer addition
  map2 <- mapdeck() %>%
    add_scatterplot(data = capitals, lon = "lon", lat = "lat")
  stopifnot(inherits(map2, "mapdeck"))
  
  cat("✓ All backward compatibility tests passed\n")
}

test_backward_compatibility()
```

### New Features Test

```{r}
# Test new multi-provider features
test_new_features <- function() {
  
  # Test 1: Provider specification
  providers <- c("mapbox", "leaflet")
  for (provider in providers) {
    if (provider %in% list_available_providers()) {
      map <- mapdeck(provider = provider)
      stopifnot(inherits(map, "mapdeck"))
      cat("✓", provider, "provider works\n")
    }
  }
  
  # Test 2: Multi-provider tokens
  set_token("test_mapbox", provider = "mapbox")
  set_token("test_leaflet", provider = "leaflet")
  tokens <- mapdeck_tokens()
  stopifnot(length(tokens) >= 2)
  cat("✓ Multi-provider tokens work\n")
  
  # Test 3: Provider switching
  if ("leaflet" %in% list_available_providers()) {
    map1 <- mapdeck(provider = "mapbox")
    map2 <- update_provider(map1, "leaflet")
    stopifnot(inherits(map2, "mapdeck"))
    cat("✓ Provider switching works\n")
  }
}

test_new_features()
```

## Troubleshooting Migration Issues

### Common Issues and Solutions

1. **"Provider not found" error**
   ```{r}
   # Check available providers
   available <- list_available_providers()
   print(available)
   
   # Use only available providers
   if ("leaflet" %in% available) {
     map <- mapdeck(provider = "leaflet")
   }
   ```

2. **Token authentication issues**
   ```{r}
   # Clear and reset tokens
   clear_tokens()
   set_token("your_new_token", provider = "mapbox")
   
   # Verify token
   token_valid <- get_token_store()$validate_token("mapbox")
   print(paste("Token valid:", token_valid))
   ```

3. **Style compatibility issues**
   ```{r}
   # Check available styles for provider
   styles <- get_available_styles("leaflet")
   print(styles)
   
   # Use compatible style
   map <- mapdeck(provider = "leaflet", style = styles[1])
   ```

4. **Coordinate system issues**
   ```{r}
   # Check coordinate system detection
   coords <- data.frame(lon = 116.4, lat = 39.9)
   crs <- detect_coordinate_system(coords)
   print(paste("Detected CRS:", crs))
   
   # Force coordinate transformation if needed
   transformed <- transform_coordinates(coords, "WGS84", "GCJ02")
   ```

## Summary

The migration to multi-provider mapdeck is designed to be seamless:

- **No breaking changes**: All existing code continues to work
- **Gradual adoption**: Add new features at your own pace  
- **Enhanced capabilities**: Access to multiple mapping providers
- **Improved flexibility**: Choose the best provider for each use case

Start by testing your existing code, then gradually explore the new multi-provider features to enhance your mapping applications.

## Next Steps

After completing your migration:

1. Explore [provider-specific examples](provider-examples.html)
2. Learn about [advanced provider features](advanced-features.html)
3. Check out [spatial sampling capabilities](spatial-sampling.html)
4. Review [performance optimization tips](performance-optimization.html)