---
title: "Provider Setup Guides"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Provider Setup Guides}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Provider Setup Guides

This guide provides comprehensive setup instructions for each mapping provider supported by mapdeck. Each provider has different requirements for authentication, configuration, and usage.

## Mapbox Provider

### Overview
Mapbox GL JS is the default provider for mapdeck, offering high-performance vector maps with 3D capabilities, custom styling, and extensive layer support.

### Authentication Setup

#### Method 1: Global Token (Recommended)
```{r}
library(mapdeck)

# Set your Mapbox access token globally
set_token("pk.your_mapbox_access_token_here", provider = "mapbox")

# Verify token is set
mapdeck_tokens()
```

#### Method 2: Environment Variable
```{r}
# Set environment variable (add to .Renviron file)
Sys.setenv(MAPBOX_TOKEN = "pk.your_mapbox_access_token_here")

# Or add to your .Renviron file:
# MAPBOX_TOKEN=pk.your_mapbox_access_token_here
```

#### Method 3: Direct Parameter
```{r}
# Pass token directly to mapdeck function
map <- mapdeck(
  token = "pk.your_mapbox_access_token_here",
  provider = "mapbox"
)
```

### Getting a Mapbox Token

1. Visit [Mapbox Account](https://account.mapbox.com/)
2. Sign up for a free account
3. Navigate to "Access tokens" section
4. Create a new token or use the default public token
5. Copy the token (starts with `pk.`)

### Basic Usage
```{r}
# Create basic Mapbox map
map <- mapdeck(
  provider = "mapbox",
  style = "streets",
  location = c(-74.006, 40.7128),  # New York
  zoom = 12,
  pitch = 45,
  bearing = 30
)
map
```

### Available Styles
```{r}
# Get available Mapbox styles
styles <- get_available_styles("mapbox")
print(styles)

# Use different styles
streets_map <- mapdeck(provider = "mapbox", style = "streets")
satellite_map <- mapdeck(provider = "mapbox", style = "satellite")
dark_map <- mapdeck(provider = "mapbox", style = "dark")
```

### Troubleshooting
- **Invalid token error**: Verify your token is correct and active
- **Rate limiting**: Check your Mapbox account usage limits
- **Style loading issues**: Ensure you have internet connectivity

---

## Leaflet Provider

### Overview
Leaflet is a popular open-source mapping library that provides tile-based maps with deck.gl overlay support. It's free to use and supports numerous tile providers.

### Setup (No Authentication Required)
```{r}
library(mapdeck)

# Leaflet doesn't require authentication tokens
map <- mapdeck(
  provider = "leaflet",
  location = c(144.9631, -37.8136),  # Melbourne
  zoom = 10
)
map
```

### Available Tile Providers
```{r}
# Get available Leaflet tile providers
providers <- get_available_styles("leaflet")
print(providers)

# Use different tile providers
osm_map <- mapdeck(provider = "leaflet", style = "streets")      # OpenStreetMap
light_map <- mapdeck(provider = "leaflet", style = "light")      # CartoDB Positron
dark_map <- mapdeck(provider = "leaflet", style = "dark")        # CartoDB DarkMatter
satellite_map <- mapdeck(provider = "leaflet", style = "satellite") # Esri WorldImagery
```

### Limitations
- **No 3D support**: Pitch and bearing are always 0
- **Limited elevation effects**: Column layers use scaled circles instead
- **Tile-based styling only**: No custom vector styles

### Best Practices
```{r}
# Use appropriate aggregation for large datasets
large_data <- data.frame(
  lon = runif(5000, 144.5, 145.5),
  lat = runif(5000, -38.2, -37.4),
  value = rnorm(5000, 100, 20)
)

# Screengrid works well with Leaflet
map <- mapdeck(provider = "leaflet", style = "light") %>%
  add_screengrid(
    data = large_data,
    lon = "lon",
    lat = "lat",
    weight = "value",
    cell_size = 50
  )
```

---

## OpenLayers Provider

### Overview
OpenLayers is a powerful open-source mapping library that supports various data sources and projections with deck.gl integration.

### Setup (No Authentication Required)
```{r}
library(mapdeck)

# OpenLayers doesn't require authentication
map <- mapdeck(
  provider = "openlayers",
  location = c(2.3522, 48.8566),  # Paris
  zoom = 12
)
map
```

### Configuration Options
```{r}
# OpenLayers with custom configuration
map <- mapdeck(
  provider = "openlayers",
  location = c(0, 0),
  zoom = 2,
  config = list(
    projection = "EPSG:3857",  # Web Mercator
    source_type = "OSM",       # OpenStreetMap tiles
    attribution = "© OpenStreetMap contributors"
  )
)
```

### Available Sources
```{r}
# Get available OpenLayers sources
sources <- get_available_styles("openlayers")
print(sources)

# Use different sources
osm_map <- mapdeck(provider = "openlayers", style = "osm")
stamen_map <- mapdeck(provider = "openlayers", style = "stamen_terrain")
```

### Advanced Features
- **Multiple projections**: Support for various coordinate systems
- **Custom sources**: Ability to add custom tile sources
- **Vector layers**: Support for vector data sources

---

## Gaode Maps Provider (高德地图)

### Overview
Gaode Maps (AutoNavi) is a popular Chinese mapping service that uses the GCJ02 coordinate system. It requires an API key for access.

### Authentication Setup

#### Getting a Gaode API Key
1. Visit [Gaode Maps Console](https://console.amap.com/)
2. Register for a developer account
3. Create a new application
4. Generate a Web API key
5. Copy the API key

#### Setting Up Authentication
```{r}
library(mapdeck)

# Set Gaode API key
set_token("your_gaode_api_key_here", provider = "gaode")

# Or use environment variable
Sys.setenv(GAODE_API_KEY = "your_gaode_api_key_here")

# Verify token is set
mapdeck_tokens()
```

### Basic Usage
```{r}
# Create Gaode map (coordinates will be auto-transformed to GCJ02)
map <- mapdeck(
  provider = "gaode",
  location = c(116.3974, 39.9093),  # Beijing (WGS84 coordinates)
  zoom = 12,
  style = "normal"
)
map
```

### Coordinate System Handling
```{r}
# Gaode uses GCJ02 coordinate system
# mapdeck automatically transforms coordinates

# WGS84 data (standard GPS coordinates)
wgs84_data <- data.frame(
  lon = c(116.3974, 121.4737),  # Beijing, Shanghai
  lat = c(39.9093, 31.2304),
  city = c("Beijing", "Shanghai")
)

# Coordinates are automatically transformed to GCJ02
map <- mapdeck(provider = "gaode") %>%
  add_scatterplot(
    data = wgs84_data,
    lon = "lon",
    lat = "lat",
    radius = 100000,
    fill_colour = "city"
  )
```

### Available Styles
```{r}
# Get available Gaode styles
styles <- get_available_styles("gaode")
print(styles)

# Use different Gaode styles
normal_map <- mapdeck(provider = "gaode", style = "normal")
satellite_map <- mapdeck(provider = "gaode", style = "satellite")
```

### Troubleshooting
- **API key errors**: Verify your key is active and has proper permissions
- **Coordinate issues**: Ensure coordinate transformation is enabled
- **Rate limiting**: Check your Gaode account usage quotas

---

## Baidu Maps Provider (百度地图)

### Overview
Baidu Maps is another major Chinese mapping service that uses the BD09 coordinate system. It requires an API key for access.

### Authentication Setup

#### Getting a Baidu API Key
1. Visit [Baidu Map Open Platform](https://lbsyun.baidu.com/)
2. Register for a developer account
3. Create a new application
4. Generate a JavaScript API key
5. Copy the API key

#### Setting Up Authentication
```{r}
library(mapdeck)

# Set Baidu API key
set_token("your_baidu_api_key_here", provider = "baidu")

# Or use environment variable
Sys.setenv(BAIDU_API_KEY = "your_baidu_api_key_here")

# Verify token is set
mapdeck_tokens()
```

### Basic Usage
```{r}
# Create Baidu map (coordinates will be auto-transformed to BD09)
map <- mapdeck(
  provider = "baidu",
  location = c(116.3974, 39.9093),  # Beijing (WGS84 coordinates)
  zoom = 12,
  style = "normal"
)
map
```

### Coordinate System Handling
```{r}
# Baidu uses BD09 coordinate system
# mapdeck automatically transforms coordinates

# WGS84 data
wgs84_data <- data.frame(
  lon = c(116.3974, 121.4737, 113.2644),  # Beijing, Shanghai, Guangzhou
  lat = c(39.9093, 31.2304, 23.1291),
  city = c("Beijing", "Shanghai", "Guangzhou")
)

# Coordinates are automatically transformed to BD09
map <- mapdeck(provider = "baidu") %>%
  add_scatterplot(
    data = wgs84_data,
    lon = "lon",
    lat = "lat",
    radius = 100000,
    fill_colour = "city"
  )
```

### Available Styles
```{r}
# Get available Baidu styles
styles <- get_available_styles("baidu")
print(styles)

# Use different Baidu styles
normal_map <- mapdeck(provider = "baidu", style = "normal")
satellite_map <- mapdeck(provider = "baidu", style = "satellite")
dark_map <- mapdeck(provider = "baidu", style = "dark")
```

---

## Provider Comparison

| Feature | Mapbox | Leaflet | OpenLayers | Gaode | Baidu |
|---------|--------|---------|------------|-------|-------|
| Authentication | Required | None | None | Required | Required |
| 3D Support | Yes | No | Limited | Yes | Yes |
| Custom Styles | Yes | Limited | Yes | Limited | Limited |
| Coordinate Systems | WGS84 | WGS84 | Multiple | GCJ02 | BD09 |
| Free Tier | Limited | Yes | Yes | Limited | Limited |
| China Access | Limited | Yes | Yes | Optimized | Optimized |

## General Troubleshooting

### Common Issues

1. **Provider not available**
   ```{r}
   # Check available providers
   providers <- list_available_providers()
   print(providers)
   ```

2. **Authentication failures**
   ```{r}
   # Validate tokens
   token_valid <- get_token_store()$validate_token("mapbox")
   print(paste("Token valid:", token_valid))
   ```

3. **Coordinate transformation issues**
   ```{r}
   # Check coordinate system detection
   coords <- data.frame(lon = 116.3974, lat = 39.9093)
   detected_crs <- detect_coordinate_system(coords)
   print(detected_crs)
   ```

4. **Performance issues**
   ```{r}
   # Check provider capabilities
   capabilities <- get_provider_capabilities("leaflet")
   print(capabilities)
   ```

### Getting Help

- Check the [mapdeck documentation](https://symbolixau.github.io/mapdeck/)
- Review provider-specific documentation
- Check GitHub issues for known problems
- Use `validate_provider_interface()` to test provider implementations

## Next Steps

After setting up your preferred provider, explore:
- [Provider-specific examples](provider-examples.html)
- [Migration guide](migration-guide.html) for switching from single to multi-provider usage
- [Advanced features](advanced-features.html) for each provider