---
title: "Provider-Specific Examples"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Provider-Specific Examples}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

# Provider-Specific Examples

This document provides comprehensive examples for each mapping provider supported by mapdeck, showcasing their unique features and capabilities.

## Overview

mapdeck supports multiple mapping providers, each with distinct characteristics:

- **Mapbox**: High-performance vector maps with 3D support
- **Leaflet**: Open-source tile-based mapping with wide provider support
- **OpenLayers**: Powerful mapping library with projection support
- **Gaode Maps**: Chinese mapping service with GCJ02 coordinate system
- **Baidu Maps**: Chinese mapping service with BD09 coordinate system

## Mapbox Provider Examples

### Basic Setup and Authentication

```{r}
library(mapdeck)

# Set Mapbox token
set_token("pk.your_mapbox_token_here", provider = "mapbox")

# Create basic Mapbox map
mapbox_basic <- mapdeck(
  provider = "mapbox",
  location = c(-74.006, 40.7128),  # New York
  zoom = 12,
  style = "streets"
)
mapbox_basic
```

### Advanced 3D Visualization

```{r}
# Load sample data
data(capitals, package = "mapdeck")

# 3D column visualization
mapbox_3d <- mapdeck(
  provider = "mapbox",
  style = "dark",
  location = c(0, 20),
  zoom = 2,
  pitch = 45,
  bearing = 30
) %>%
  add_column(
    data = capitals,
    lon = "lon",
    lat = "lat",
    elevation = "population",
    fill_colour = "country",
    radius = 50000,
    tooltip = c("capital_city", "country", "population")
  )
mapbox_3d
```

### Custom Mapbox Styles

```{r}
# Available Mapbox styles
mapbox_styles <- get_available_styles("mapbox")
print(mapbox_styles)

# Use different styles
satellite_map <- mapdeck(
  provider = "mapbox",
  style = "satellite",
  location = c(-122.4194, 37.7749),  # San Francisco
  zoom = 14
)

# Custom style URL
custom_style <- mapdeck(
  provider = "mapbox",
  style = "mapbox://styles/mapbox/navigation-day-v1",
  location = c(2.3522, 48.8566),  # Paris
  zoom = 12
)
```

### Animated Layers

```{r}
# Animated arc connections
mapbox_animated <- mapdeck(
  provider = "mapbox",
  style = "dark"
) %>%
  add_animated_arc(
    data = capitals,
    origin = c("lon", "lat"),
    destination = c("lon", "lat"),
    stroke_from = "country",
    stroke_to = "country",
    animation_speed = 2000,
    stroke_width = 4
  )
mapbox_animated
```

## Leaflet Provider Examples

### Basic Tile Providers

```{r}
# Leaflet doesn't require authentication
leaflet_basic <- mapdeck(
  provider = "leaflet",
  location = c(144.9631, -37.8136),  # Melbourne
  zoom = 10,
  style = "streets"  # OpenStreetMap
)
leaflet_basic

# Different tile providers
cartodb_light <- mapdeck(
  provider = "leaflet",
  style = "light",  # CartoDB Positron
  location = c(-74.006, 40.7128),
  zoom = 12
)

stamen_watercolor <- mapdeck(
  provider = "leaflet",
  style = "watercolor",  # Stamen Watercolor
  location = c(2.3522, 48.8566),
  zoom = 12
)
```

### Leaflet Layer Compatibility

```{r}
data(melbourne, package = "mapdeck")

# Multiple layers on Leaflet
leaflet_multi <- mapdeck(
  provider = "leaflet",
  style = "streets",
  location = c(144.9631, -37.8136),
  zoom = 10
) %>%
  add_polygon(
    data = melbourne,
    fill_colour = "SA2_NAME",
    fill_opacity = 0.5,
    stroke_colour = "#FFFFFF",
    stroke_width = 20,
    layer_id = "boundaries"
  ) %>%
  add_scatterplot(
    data = capitals,
    lon = "lon",
    lat = "lat",
    radius = 30000,
    fill_colour = "#FF4444",
    layer_id = "cities"
  ) %>%
  add_text(
    data = capitals,
    lon = "lon",
    lat = "lat",
    text = "capital_city",
    size = 12,
    colour = "#000000",
    layer_id = "labels"
  )
leaflet_multi
```

### Performance Optimization for Leaflet

```{r}
# Large dataset with aggregation
large_data <- data.frame(
  lon = runif(5000, 144.5, 145.5),
  lat = runif(5000, -38.2, -37.4),
  value = rnorm(5000, 100, 20)
)

# Screengrid aggregation works well with Leaflet
leaflet_screengrid <- mapdeck(
  provider = "leaflet",
  style = "light",
  location = c(145, -37.8),
  zoom = 9
) %>%
  add_screengrid(
    data = large_data,
    lon = "lon",
    lat = "lat",
    weight = "value",
    cell_size = 50,
    opacity = 0.8
  )
leaflet_screengrid
```

## OpenLayers Provider Examples

### Multiple Data Sources

```{r}
# Basic OpenLayers map
openlayers_basic <- mapdeck(
  provider = "openlayers",
  location = c(2.3522, 48.8566),  # Paris
  zoom = 12,
  style = "osm"  # OpenStreetMap
)
openlayers_basic

# Different tile sources
stamen_terrain <- mapdeck(
  provider = "openlayers",
  style = "stamen_terrain",
  location = c(-122.4194, 37.7749),  # San Francisco
  zoom = 10
)

cartodb_dark <- mapdeck(
  provider = "openlayers",
  style = "cartodb_dark",
  location = c(151.2093, -33.8688),  # Sydney
  zoom = 11
)
```

### Custom Projections

```{r}
# Web Mercator projection (default)
web_mercator <- mapdeck(
  provider = "openlayers",
  location = c(0, 0),
  zoom = 2,
  config = list(projection = "EPSG:3857")
)

# Geographic projection
geographic <- mapdeck(
  provider = "openlayers",
  location = c(0, 0),
  zoom = 2,
  config = list(projection = "EPSG:4326")
)
```

### Advanced OpenLayers Features

```{r}
# Multiple layers with different sources
openlayers_advanced <- mapdeck(
  provider = "openlayers",
  style = "cartodb_light",
  location = c(144.9631, -37.8136),
  zoom = 8
) %>%
  add_scatterplot(
    data = capitals,
    lon = "lon",
    lat = "lat",
    radius = 50000,
    fill_colour = "#FF6B6B",
    layer_id = "points"
  ) %>%
  add_polygon(
    data = melbourne,
    fill_colour = "SA2_NAME",
    fill_opacity = 0.3,
    stroke_colour = "#4ECDC4",
    stroke_width = 10,
    layer_id = "polygons"
  )
openlayers_advanced
```

## Chinese Providers Examples

### Gaode Maps (高德地图)

```{r}
# Set up Gaode authentication
set_token("your_gaode_api_key_here", provider = "gaode")

# Chinese cities data (WGS84 coordinates - automatically transformed)
chinese_cities <- data.frame(
  city_cn = c("北京", "上海", "广州", "深圳", "成都"),
  city_en = c("Beijing", "Shanghai", "Guangzhou", "Shenzhen", "Chengdu"),
  lon = c(116.3974, 121.4737, 113.2644, 114.0579, 104.0668),
  lat = c(39.9093, 31.2304, 23.1291, 22.5431, 30.5728),
  population = c(21540000, 24870000, 15300000, 12530000, 16330000)
)

# Gaode map with Chinese cities
gaode_cities <- mapdeck(
  provider = "gaode",
  location = c(116.3974, 39.9093),  # Beijing
  zoom = 5,
  style = "normal"
) %>%
  add_scatterplot(
    data = chinese_cities,
    lon = "lon",
    lat = "lat",
    radius = "population",
    fill_colour = "city_cn",
    tooltip = c("city_cn", "city_en", "population"),
    radius_min_pixels = 10,
    radius_max_pixels = 100
  ) %>%
  add_text(
    data = chinese_cities,
    lon = "lon",
    lat = "lat",
    text = "city_cn",  # Chinese city names
    size = 16,
    colour = "#2C3E50"
  )
gaode_cities
```

### Baidu Maps (百度地图)

```{r}
# Set up Baidu authentication
set_token("your_baidu_api_key_here", provider = "baidu")

# Same data, different provider (coordinates auto-transformed to BD09)
baidu_cities <- mapdeck(
  provider = "baidu",
  location = c(116.3974, 39.9093),  # Beijing
  zoom = 5,
  style = "normal"
) %>%
  add_scatterplot(
    data = chinese_cities,
    lon = "lon",
    lat = "lat",
    radius = "population",
    fill_colour = "city_cn",
    tooltip = c("city_cn", "city_en", "population"),
    radius_min_pixels = 10,
    radius_max_pixels = 100
  )
baidu_cities
```

### Coordinate System Handling

```{r}
# Manual coordinate transformation examples
wgs84_coords <- data.frame(
  lon = c(116.3974, 121.4737),
  lat = c(39.9093, 31.2304)
)

# Transform to GCJ02 (Gaode)
gcj02_coords <- transform_coordinates(wgs84_coords, "WGS84", "GCJ02")
print("GCJ02 coordinates for Gaode:")
print(gcj02_coords)

# Transform to BD09 (Baidu)
bd09_coords <- transform_coordinates(wgs84_coords, "WGS84", "BD09")
print("BD09 coordinates for Baidu:")
print(bd09_coords)

# Detect coordinate system
detected_crs <- detect_coordinate_system(wgs84_coords)
print(paste("Detected coordinate system:", detected_crs))
```

## Cross-Provider Comparisons

### Same Visualization, Different Providers

```{r}
# Function to create the same visualization across providers
create_cross_provider_comparison <- function(data) {
  
  base_config <- list(
    location = c(0, 20),
    zoom = 2
  )
  
  # Mapbox version
  mapbox_map <- mapdeck(provider = "mapbox", !!!base_config) %>%
    add_scatterplot(
      data = data,
      lon = "lon",
      lat = "lat",
      radius = 100000,
      fill_colour = "country",
      tooltip = c("capital_city", "country")
    )
  
  # Leaflet version
  leaflet_map <- mapdeck(provider = "leaflet", !!!base_config) %>%
    add_scatterplot(
      data = data,
      lon = "lon",
      lat = "lat",
      radius = 100000,
      fill_colour = "country",
      tooltip = c("capital_city", "country")
    )
  
  # OpenLayers version
  openlayers_map <- mapdeck(provider = "openlayers", !!!base_config) %>%
    add_scatterplot(
      data = data,
      lon = "lon",
      lat = "lat",
      radius = 100000,
      fill_colour = "country",
      tooltip = c("capital_city", "country")
    )
  
  return(list(
    mapbox = mapbox_map,
    leaflet = leaflet_map,
    openlayers = openlayers_map
  ))
}

# Create comparison maps
comparison_maps <- create_cross_provider_comparison(capitals)
```

### Provider Switching Examples

```{r}
# Start with one provider
original_map <- mapdeck(
  provider = "mapbox",
  style = "streets",
  location = c(-74.006, 40.7128),
  zoom = 12
) %>%
  add_scatterplot(
    data = capitals,
    lon = "lon",
    lat = "lat",
    radius = 50000,
    fill_colour = "country"
  )

# Switch to different providers while preserving layers
leaflet_switched <- update_provider(original_map, "leaflet")
openlayers_switched <- update_provider(original_map, "openlayers")

# Check compatibility before switching
compatibility <- check_provider_compatibility("mapbox", "leaflet")
print("Compatibility check:")
print(compatibility)
```

## Performance Comparisons

### Large Dataset Handling

```{r}
# Create large dataset for performance testing
large_dataset <- data.frame(
  lon = runif(10000, -180, 180),
  lat = runif(10000, -85, 85),
  value = rnorm(10000, 100, 20),
  category = sample(LETTERS[1:5], 10000, replace = TRUE)
)

# Performance test function
test_provider_performance <- function(provider, data) {
  start_time <- Sys.time()
  
  map <- mapdeck(provider = provider, location = c(0, 0), zoom = 2) %>%
    add_screengrid(
      data = data,
      lon = "lon",
      lat = "lat",
      weight = "value",
      cell_size = 50
    )
  
  end_time <- Sys.time()
  return(as.numeric(end_time - start_time))
}

# Test different providers (if available)
providers_to_test <- c("mapbox", "leaflet", "openlayers")
performance_results <- list()

for (provider in providers_to_test) {
  if (provider %in% list_available_providers()) {
    performance_results[[provider]] <- test_provider_performance(provider, large_dataset)
    cat(provider, "performance:", round(performance_results[[provider]], 3), "seconds\n")
  }
}
```

## Specialized Use Cases

### Scientific Visualization

```{r}
# Scientific data visualization with different providers
scientific_data <- data.frame(
  lon = runif(1000, -180, 180),
  lat = runif(1000, -85, 85),
  temperature = rnorm(1000, 15, 10),
  pressure = rnorm(1000, 1013, 50),
  humidity = runif(1000, 0, 100)
)

# Mapbox for high-quality scientific visualization
scientific_mapbox <- mapdeck(
  provider = "mapbox",
  style = "satellite",
  location = c(0, 0),
  zoom = 2
) %>%
  add_scatterplot(
    data = scientific_data,
    lon = "lon",
    lat = "lat",
    radius = "temperature",
    fill_colour = "pressure",
    tooltip = c("temperature", "pressure", "humidity")
  )

# Leaflet for open-source scientific applications
scientific_leaflet <- mapdeck(
  provider = "leaflet",
  style = "terrain",
  location = c(0, 0),
  zoom = 2
) %>%
  add_heatmap(
    data = scientific_data,
    lon = "lon",
    lat = "lat",
    weight = "temperature",
    radius_pixels = 60
  )
```

### Business Intelligence Dashboards

```{r}
# Business data
business_data <- data.frame(
  city = c("New York", "London", "Tokyo", "Sydney", "Toronto"),
  lon = c(-74.006, -0.1276, 139.6917, 151.2093, -79.3832),
  lat = c(40.7128, 51.5074, 35.6895, -33.8688, 43.6532),
  revenue = c(1200000, 980000, 1500000, 750000, 650000),
  employees = c(450, 320, 580, 280, 220),
  growth = c(15.2, 8.7, 22.1, 12.3, 9.8)
)

# Executive dashboard with Mapbox
executive_dashboard <- mapdeck(
  provider = "mapbox",
  style = "dark",
  location = c(0, 30),
  zoom = 2
) %>%
  add_column(
    data = business_data,
    lon = "lon",
    lat = "lat",
    elevation = "revenue",
    fill_colour = "growth",
    radius = 100000,
    tooltip = c("city", "revenue", "employees", "growth")
  ) %>%
  add_text(
    data = business_data,
    lon = "lon",
    lat = "lat",
    text = "city",
    size = 14,
    colour = "#FFFFFF"
  )
executive_dashboard
```

### Educational Applications

```{r}
# Educational geography data
world_capitals <- data.frame(
  capital = c("Washington D.C.", "London", "Paris", "Berlin", "Rome"),
  country = c("USA", "UK", "France", "Germany", "Italy"),
  lon = c(-77.0369, -0.1276, 2.3522, 13.4050, 12.4964),
  lat = c(38.9072, 51.5074, 48.8566, 52.5200, 41.9028),
  population = c(705749, 8982000, 2161000, 3669491, 2873000),
  founded = c(1790, 43, -250, 1237, -753)
)

# Educational map with Leaflet (free for educational use)
educational_map <- mapdeck(
  provider = "leaflet",
  style = "streets",
  location = c(10, 50),
  zoom = 3
) %>%
  add_scatterplot(
    data = world_capitals,
    lon = "lon",
    lat = "lat",
    radius = "population",
    fill_colour = "country",
    tooltip = c("capital", "country", "population", "founded"),
    radius_min_pixels = 10,
    radius_max_pixels = 50
  ) %>%
  add_text(
    data = world_capitals,
    lon = "lon",
    lat = "lat",
    text = "capital",
    size = 12,
    colour = "#2C3E50"
  )
educational_map
```

## Integration Examples

### Shiny Applications

```{r}
if (FALSE) {  # Shiny app example
  library(shiny)
  
  ui <- fluidPage(
    titlePanel("Multi-Provider Map Comparison"),
    sidebarLayout(
      sidebarPanel(
        selectInput("provider", "Select Provider:",
                   choices = list_available_providers(),
                   selected = "mapbox"),
        selectInput("style", "Map Style:",
                   choices = NULL),  # Populated based on provider
        sliderInput("zoom", "Zoom Level:", min = 1, max = 18, value = 10),
        checkboxInput("show_3d", "Enable 3D (if supported)", value = FALSE)
      ),
      mainPanel(
        mapdeckOutput("map", height = "600px"),
        verbatimTextOutput("provider_info")
      )
    )
  )
  
  server <- function(input, output, session) {
    
    # Update style choices based on provider
    observe({
      styles <- get_available_styles(input$provider)
      updateSelectInput(session, "style", choices = styles, selected = styles[1])
    })
    
    # Create reactive map
    output$map <- renderMapdeck({
      
      # Get provider capabilities
      capabilities <- get_provider_capabilities(input$provider)
      
      # Adjust 3D settings based on provider support
      pitch <- if (input$show_3d && capabilities$supports_3d) 45 else 0
      bearing <- if (input$show_3d && capabilities$supports_3d) 30 else 0
      
      map <- mapdeck(
        provider = input$provider,
        style = input$style,
        zoom = input$zoom,
        pitch = pitch,
        bearing = bearing,
        location = c(0, 20)
      ) %>%
        add_scatterplot(
          data = capitals,
          lon = "lon",
          lat = "lat",
          radius = 100000,
          fill_colour = "country"
        )
      
      map
    })
    
    # Display provider information
    output$provider_info <- renderText({
      capabilities <- get_provider_capabilities(input$provider)
      paste(
        "Provider:", input$provider,
        "\nSupports 3D:", capabilities$supports_3d,
        "\nCoordinate System:", capabilities$coordinate_system,
        "\nAuthentication Required:", capabilities$requires_auth
      )
    })
  }
  
  shinyApp(ui = ui, server = server)
}
```

## Best Practices Summary

### Provider Selection Guidelines

1. **Mapbox**: Choose for high-quality 3D visualizations and custom styling
2. **Leaflet**: Choose for open-source projects and educational use
3. **OpenLayers**: Choose for advanced projection support and custom data sources
4. **Gaode**: Choose for applications targeting Chinese users
5. **Baidu**: Choose for China-focused applications with local optimization

### Performance Optimization

- Use aggregation layers (screengrid, hexagon, grid) for large datasets
- Choose appropriate providers based on your performance requirements
- Test different providers with your specific data and use case
- Consider network connectivity and geographic location of users

### Development Workflow

1. Start with a single provider for prototyping
2. Test your visualization across multiple providers
3. Use provider switching to compare performance and appearance
4. Implement fallback providers for robustness
5. Document provider-specific requirements and limitations

This comprehensive guide provides the foundation for leveraging the full power of mapdeck's multi-provider architecture in your applications.